Appendix 1: Preregistered analysis with simulated data

### 2025-04-07 preregistration.Rmd
reorganise models

@@

Having a donor self-identity is considered a key motivational factor for sustained blood donation behaviour (Callero, 1985; Ferguson et al., 2020). Within Identity Theory, when a new donor first gives blood they develop a role-identity (I am a blood donor). As a novice donor continues to donate, this role-identity is validated, and thus blood donation becomes part of a donor’s self-identity (Being a blood donor is an important part of who I am; Grube & Piliavin, 2000; Masser, White, Hyde, Terry, & Robinson, 2009). However, despite the link between donor self-identity and donor retention, there are currently no interventions which can foster a sense of identity in donors that do not rely exclusively on repeat donation behaviour. 
Psychological ownership (this is mine) is the sense of ownership experienced over both material (e.g., blood collection agency) and abstract immaterial (e.g., blood donation behaviour) possessions (Pierce et al., 2001). The ownership volunteers experience over a non-profit organisation has been linked to sustained volunteering behaviour (Ainsworth, 2020). Further, psychological ownership is theoretically associated with identity as it has been proposed that through possession identity is established and maintained (Dittmar, 1992; Pierce et al., 2001). 
In an Australian donor context, we have provided initial support for the idea that psychological ownership experienced towards a blood collection agency both increases with past donation experience and is positively associated with blood donor self-identity (Edwards et al., 2023). Further, in a series of semi-structured interviews we have shown that Australian blood donors primarily expressed ownership over the blood donation process (role behaviour), not the blood collection agency (organisation). As such, interventions that enhance the amount of psychological ownership donors experience over their donation have the potential to impact donor retention.

FRCBS

@@

model selection and then ci

@@

All donors answering to question

@@

>>>>

TODO:

- references

@@

Having a donor self-identity is considered a key motivational factor for sustained blood donation behaviour (Callero, 1985; Ferguson et al., 2020). Within Identity Theory, when a new donor first gives blood they develop a role-identity (I am a blood donor). As a novice donor continues to donate, this role-identity is validated, and thus blood donation becomes part of a donor’s self-identity (Being a blood donor is an important part of who I am; Grube & Piliavin, 2000; Masser, White, Hyde, Terry, & Robinson, 2009). However, despite the link between donor self-identity and donor retention, there are currently no interventions which can foster a sense of identity in donors that do not rely exclusively on repeat donation behaviour. 
Psychological ownership (this is mine) is the sense of ownership experienced over both material (e.g., blood collection agency) and abstract immaterial (e.g., blood donation behaviour) possessions (Pierce et al., 2001). The ownership volunteers experience over a non-profit organisation has been linked to sustained volunteering behaviour (Ainsworth, 2020). Further, psychological ownership is theoretically associated with identity as it has been proposed that through possession identity is established and maintained (Dittmar, 1992; Pierce et al., 2001). 
In an Australian donor context, we have provided initial support for the idea that psychological ownership experienced towards a blood collection agency both increases with past donation experience and is positively associated with blood donor self-identity (Edwards et al., 2023). Further, in a series of semi-structured interviews we have shown that Australian blood donors primarily expressed ownership over the blood donation process (role behaviour), not the blood collection agency (organisation). As such, interventions that enhance the amount of psychological ownership donors experience over their donation have the potential to impact donor retention. 
In this study we try to repeat results from Edwards et al., 2023 in Finland by collecting similar questionnaire data and analysing them with a structural equation model. We include pre-registered R analysis code for direct repetition of Edwards et al., 2023 in Appendix 1.

@@

We will model the hypothesis through alternative models:

a. Sum score model with robust estimator

b. Sum score model with bootstrapping

c. Latent factor model with robust estimator

d. Latent factor model with boostrapping

Among

@@

All donors answering to question de01 other than "man" or "women" will be excluded as donation policies depend on sex. All donors answering to question bd01 "no" will be excluded as we are interested in person eligible to donate.

@@

The survey platform only reports completed questionnaires and all questions are obligatory. Hence there will be no missing data.

### 2025-04-07 preregistration_code.Rmd
reorganise models

---
title: "Supplementary file 1: Preregistered analysis"
author: "Abigail Edwards, Lauri Hietajärvi and Mikko Arvas"
date: "`r Sys.time()`"
output:
  pdf_document:
    toc: yes

@@

@@

#lavaan does not understand empty factor levels
data.fit <- data %>%
  #lavaan cannot deal with unordered factor with more then 2 levels
  filter(de01 == 'man' | de01 == "women") %>% # exclusion
  filter(bd01 != 'no') %>% # exclusion
  mutate(
    across(isfactor, ~ fct_drop(.x)
    ) %>% 
      mutate(
        bd04 = as.numeric(bd04), # lavaan cannot handle dates
        bd04 = bd04 / 10^-(2 - str_length(as.numeric(max(bd04)))) # https://groups.google.com/g/lavaan/c/r7w-4HHg5R0 
        #"Your underweight variance seems to be much larger than other variances.  Try dividing it by 10 or 100 to make the SDs more similar across modeled variables." 
        # But this does not seem to necessarily help.
      )
  ) %>% 
  #Original paper used sum scores instead of CFA. Let's add them for comparison
  mutate(
    ss_psycown = as.numeric(po01) + as.numeric(po02) + as.numeric(po03),
    ss_selfid = as.numeric(si01) + as.numeric(si02) + as.numeric(si03),
    ss_intention = as.numeric(it01) + as.numeric(it02) + as.numeric(it03),
    #donation history is on diffent scales, normalise them before summing.
    #others should be positively correlated, but bd04 negatively
    ss_blooddonor = scale(as.numeric(bd02)) 
    + scale(as.numeric(bd01)) 
    + scale(bd03)  
    -scale(bd04) # do we want to include the DATE OF LAST DONATION?

@@

### Sum score model

@@

### SEM

@@

### Sum score with bootstrapping

fit.po03ss <- sem(po02ss.model,  
                    estimator = "ML", # bootstrapping does not work with estimators designed for factors
                    data = data.fit2, # data that is completely numeric
                bootstrap = nBoots,
                se = "boot",
                auto.fix.first = FALSE,
                std.lv = TRUE
                )
models[['fit.po03ss']] <- fit.po03ss
}

### SEM with bootstrapping

@@

### SEM with Bayes

```{r}
po02b.model <- ' 
#measurement model
psycown =~ po01 + po02 + po03
selfid =~ si01 + si02 + si03
intention =~ it01 + it02 + it03
blooddonor =~ bd02 + bd03 +bd04

# #mediation
 psycown ~ a1 *  blooddonor
 selfid ~ a2 * blooddonor + d21 * psycown
# 
# #regression
 intention ~  de01 + de02 + ie01 +  cp * blooddonor + b1 * psycown + b2 * selfid
ind_eff := a1* d21 * b2
'

```{r}
# This does not work!
# #blavaan
#tmp <- as.data.frame(data.fit2)
 # fit.po04 <- blavaan(po02.model,  
 #         #data = data.fit2,
 #         data = tmp,
 #         auto.fix.first = FALSE,
 #         std.lv = TRUE
 #)
#Error in validObject(.Object) : 
#invalid class “Fit” object: invalid object for slot "fx.group" in class "Fit": got class "NULL", should be or extend class "numeric"
#We will try to fix this with real data
```

# #This works!
# data(HolzingerSwineford1939, package = "lavaan")
# # The Holzinger and Swineford (1939) example
# HS.model <- ' visual  =~ x1 + x2 + x3
#               textual =~ x4 + x5 + x6
#               speed   =~ x7 + x8 + x9 '
# 
# fit <- blavaan(HS.model, data = HolzingerSwineford1939,
#                auto.var = TRUE, auto.fix.first = TRUE,
#                auto.cov.lv.x = TRUE)
# summary(fit)
# coef(fit)

>>>>

---
title: "Appendix 1: Preregistered analysis with simulated data"
author: "Abigail Edwards, Lauri Hietajärvi and Mikko Arvas"
date: "`r Sys.time()`"
output:
  pdf_document:
    toc: yes

@@

 -report effects of exclusions

@@

#lavaan does not understand empty factor levels
data.fit <- data %>%
  #lavaan cannot deal with unordered factor with more then 2 levels
  filter(de01 == 'man' | de01 == "women") %>% # EXCLUSION
  filter(bd01 != 'no') %>% # EXCLUSION
  mutate(
    across(isfactor, ~ fct_drop(.x)
    ) %>% 
      mutate(
        bd04 = as.numeric(bd04), # lavaan cannot handle dates
        bd04 = bd04 / 10^-(2 - str_length(as.numeric(max(bd04)))) # https://groups.google.com/g/lavaan/c/r7w-4HHg5R0 
        #"Your underweight variance seems to be much larger than other variances.  Try dividing it by 10 or 100 to make the SDs more similar across modeled variables." 
        # But this does not seem to necessarily help.
      )
  ) %>% 
  #Original paper used sum scores instead of CFA. Let's add them for comparison
  mutate(
    ss_psycown = as.numeric(po01) + as.numeric(po02) + as.numeric(po03),
    ss_selfid = as.numeric(si01) + as.numeric(si02) + as.numeric(si03),
    ss_intention = as.numeric(it01) + as.numeric(it02) + as.numeric(it03),
    #donation history is on diffent scales, normalise them before summing.
    #others should be positively correlated, but bd04 negatively
    ss_blooddonor = scale(as.numeric(bd02)) 
    + scale(as.numeric(bd01)) 
    + scale(bd03)  
    -scale(bd04) # do we want to include the DATE OF LAST DONATION?

@@

### a) Sum score model with robust estimator

@@

###  b) Sum score model with bootstrapping

fit.po03ss <- sem(po02ss.model,  
                    estimator = "ML", # 
                    data = data.fit2, # data that is completely numeric
                bootstrap = nBoots,
                se = "boot",
                auto.fix.first = FALSE,
                std.lv = TRUE
                )
models[['fit.po03ss']] <- fit.po03ss
}

### c) Latent factor model with robust estimator

@@

### d) Latent factor model with boostrapping

@@
